---
title: "Decision trees"
author: "James Karebe"
date: "2022-10-27"
output: html_document
---

```{r}
setwd("D:School/3/Applied Analytics")
credit <- read.csv("D:/School/3/Applied Analytics/Credit_Data.csv",stringsAsFactors = TRUE)
```


## Exploring data

```{r}
str(credit)
```

Exploring variables of concern and their distributions.

```{r}
table(credit$credit_history)
table(credit$savings)
table(credit$job)
table(credit$duration_in_month)
```

```{r}
summary(credit$duration_in_month)
summary(credit$savings)
```

Duration in months is skewed to the right.

Changing the dependent variable labels;

```{r}
credit$default <- factor(credit$default, levels = c("0","1"), labels = c("No","Yes"))
table(credit$default)
```

To randomize the order of the data

```{r}
set.seed(12345)
credit_r <- credit[order(runif(1000)), ]
```

Change the data from reading as levels to read as names

```{r}
levels(credit_r$savings)<-make.names(levels(credit_r$savings))
levels(credit_r$account_check_status)<-make.names(levels(credit_r$account_check_status))
levels(credit_r$present_emp_since)<-make.names(levels(credit_r$present_emp_since))
levels(credit_r$personal_status_sex)<-make.names(levels(credit_r$personal_status_sex))
levels(credit_r$property)<-make.names(levels(credit_r$property))
levels(credit_r$credit_history)<-make.names(levels(credit_r$credit_history))
levels(credit_r$foreign_worker)<-make.names(levels(credit_r$foreign_worker))
levels(credit_r$housing)<-make.names(levels(credit_r$housing))
levels(credit_r$job)<-make.names(levels(credit_r$job))
levels(credit_r$other_debtors)<-make.names(levels(credit_r$other_debtors))
levels(credit_r$other_installment_plans)<-make.names(levels(credit_r$other_installment_plans))
levels(credit_r$purpose)<-make.names(levels(credit_r$purpose))
levels(credit_r$telephone)<-make.names(levels(credit_r$telephone))
```

Explore the randomized dataset;

```{r}
summary(credit_r$duration_in_month)
summary(credit$duration_in_month)
```

Comparing the first few values;

```{r}
head(credit$credit_amount)
head(credit_r$credit_amount)
```


## Training the data

Split the data for training and testing *90/10*

```{r}
credit_train <- credit_r[1:900, ]
credit_test <- credit_r[901:1000, ]

prop.table(table(credit_test$default))
prop.table(table(credit_train$default))
```

It is a fairly equal split

Implementing a c5.0 algorithm

```{r}
library(C50)
credit_model <- C5.0(credit_train[-1],credit_train$default)
credit_model
summary(credit_model)
```

The model incorrectly predicts 127 cases (14.1%)

TP = 590, FP = 85, TN = 183, FN = 42


## Evaluating model performance

Predicting the model to our test data set

```{r}
credit_pred <- predict(credit_model, credit_test)
library(gmodels)
CrossTable(credit_test$default,credit_pred, 
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('actual default', 'predicted default'))
```

The model correctly predicts 75% and incorrect 25%. However false negatives need to be as minimal as possible (higher false negatives are costly).


## Improving model performance

Using adaptive boosting

```{r}
credit_boost10 <- C5.0(credit_train[-1], credit_train$default, trials = 10)
credit_boost10
summary(credit_boost10)
```
There are 30 incorrect predictions (3.33% error)

```{r}
credit_boost_pred10 <- predict(credit_boost10, credit_test)
CrossTable(credit_test$default,credit_boost_pred10, 
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('actual default', 'predicted default'))
```
The model correctly predicts 79%. It also has reduced false positive to 16

